\section[PROJECT SCHEDULE]{Project Schedule}

\subsection{Phase I: Proposal and Planning (January -- February 2026)}

\textbf{Objectives:}
\begin{itemize}
    \item Finalize system architecture and refine research scope.
    \item Consolidate focused literature review.
    \item Prepare dataset acquisition and preprocessing design.
\end{itemize}

\textbf{Milestones:}
\begin{itemize}
    \item Assignment of Supervisor -- January 2, 2026
    \item Submission of Research Proposal -- February 15, 2026
    \item Proposal Defense Presentation -- February 25, 2026
\end{itemize}

\textbf{Deliverable:} Approved and refined research proposal.

\subsection{Phase II: Core Framework Development (Mode II) (March -- May 2026)}

This phase focuses exclusively on implementing the lightweight projected 
privacy framework without cryptographic integration.

\textbf{Objectives:}
\begin{itemize}
    \item Implement multimodal feature extraction (Keystroke + Mouse).
    \item Develop sliding window fusion mechanism.
    \item Implement Johnson--Lindenstrauss (JL) projection layer.
    \item Train Deep SVDD model in plaintext.
    \item Evaluate baseline biometric performance (EER, FAR, FRR).
\end{itemize}

\textbf{Milestones:}
\begin{itemize}
    \item Submission of Introduction and Literature Review -- May 17, 2026
    \item Mid Review Presentation -- May 27, 2026
\end{itemize}

\textbf{Deliverable:} Functional Mode II prototype with preliminary results.

\subsection{Phase III: Performance Optimization and Validation (June -- July 2026)}

\textbf{Objectives:}
\begin{itemize}
    \item Optimize projection dimension $k$.
    \item Measure CPU-based inference latency.
    \item Conduct $k$-fold cross-validation.
    \item Perform zero-effort impostor testing.
    \item Compare projected vs non-projected baseline models.
\end{itemize}

\textbf{Deliverable:} Validated lightweight framework with 
accuracy--latency trade-off analysis.

\subsection{Phase IV: Cryptographic Integration (Mode I -- Limited Scope) (August 2026)}

To ensure feasibility, the cryptographic module is implemented as a 
controlled proof-of-concept with reduced scope.

\textbf{Scope Limitation Strategy:}
\begin{itemize}
    \item Encrypt only JL-projected embeddings.
    \item Perform encrypted squared Euclidean distance computation only.
    \item Deep SVDD training remains in plaintext.
    \item No full encrypted neural network training.
\end{itemize}

\textbf{Milestone:}
\begin{itemize}
    \item Submission of Methodology Chapter -- August 9, 2026
\end{itemize}

\textbf{Deliverable:} Proof-of-concept encrypted inference with 
encryption overhead analysis.

\subsection{Phase V: Security Evaluation and Finalization (September -- October 2026)}

\textbf{Objectives:}
\begin{itemize}
    \item Conduct reconstruction attack evaluation (JL mode).
    \item Measure mutual information leakage.
    \item Perform final privacy--latency--accuracy comparison.
    \item Complete thesis writing and system documentation.
\end{itemize}

\textbf{Milestones:}
\begin{itemize}
    \item Submission of Final Project Report -- September 27, 2026
    \item Final Presentation and Demonstration -- October 19--20, 2026
\end{itemize}

\textbf{Final Deliverables:}
\begin{itemize}
    \item Complete research thesis.
    \item Functional Mode II system.
    \item Mode I encrypted inference proof-of-concept.
\end{itemize}



